---
title       : Permeability Predictor
subtitle    : a product of D.E., Inc.
author      : Jeff Wolbach
job         : Overeducated Layabout
logo        : Heinz_Logo.png
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## Description of Problem

- Oral bioavailability is an important consideration in drug development

- High oral bioavailability necessary for dosing as tablets

- Tablet dosing leads to much stronger patient compliance and less-expensive treatment.

- It is very economically advantageous to fail compounds with poor oral bioavailability early in development


--- &twocol w1:40% w2:60%

## Available Measures
*** =left
- Ideal measure would be fractional drug absorbance in humans.  But ... <i>in vivo</i> assays are difficult and expensive
- <i>in vitro</i> assays are much more cost-effective and parallelizable
- Assays have been shown to correlate well with fractional drug absorbance.

*** =right
```{r,echo=FALSE, results='asis'}
logPlot<-seq(-3, 2, by=0.04)
PlotXs<-10^(logPlot)
PlotYs<-100/(1+(1/(sqrt(10)*PlotXs)))
par(bg="white", pin=c(5,5))
plot(logPlot,PlotYs,xlab="log10 Permability",ylab="% Frac Absorbed, Human", type="n")
x<-par("usr")
xRed<-(2*x[1]+x[2])/3
rect(x[1],x[3],xRed,x[4],col="firebrick3")
xGrn<-(x[1]+2*x[2])/3
rect(xGrn,x[3],x[2],x[4],col="lightgreen")
points(logPlot,PlotYs, type="l", col="darkblue", lwd=3)
```

--- .class #id

## Even Cheaper Than Experiment

- We can model the in vitro permeability using a few physical parameters of our molecules

log P(apparent) = -1.64 + 0.85 * log K(O/W) - 0.025 * TPSA - 0.27 * (pKa - pH)

K(O/w) = octanol-water partition coefficient

TPSA = Total polar surface area of molecule

pKa = -base-10 logarithm of acidity constant for molecule

pH = -base-10 logarithm of H+ activity in assay solution

(pKa - pH) = likelihood molecule is charged


--- &twocol w1:40% w2:60%

## Utility of Application

*** =left
```{r,echo=FALSE}
pAppModel<-function(pH, pKa, logP, TPSA) {
  -1.64 + 0.85*logP - 0.025*TPSA - 0.27*(pKa-pH)
}
logPlot<-seq(-3, 2, by=0.04)
PlotXs<-10^(logPlot)
PlotYs<-100/(1+(1/(sqrt(10)*PlotXs)))
par(bg="lightgray", pin=c(5,4))
plot(logPlot,PlotYs,xlab="log10 Permability",ylab="% Frac Absorbed, Human", type="n")
x<-par("usr")
xRed<-(2*x[1]+x[2])/3
rect(x[1],x[3],xRed,x[4],col="firebrick3")
xGrn<-(x[1]+2*x[2])/3
rect(xGrn,x[3],x[2],x[4],col="lightgreen")
points(logPlot,PlotYs, type="l", col="darkblue", lwd=3)
xCalc<-pAppModel(6.5, 8.2, 3.1, 40)
xCalc2<-10^xCalc
yCalc<-100/(1+(1/(sqrt(10)*xCalc2)))
points(xCalc, yCalc, type="p", pch=16, col="yellow", cex=4.5)
```

*** =right
- Properties are easily calculated from a webform
- Analyst can visualize the permeability of their compound, even before synthesis.
- Allows analysts to concentrate on making compounds that will not fail early

- Next generation of app will allow user to draw molecule to create a fully-automated workflow.





